---
title: "DS 6306 Project 1"
author: "William Jones & Christian Orji"
date: "2023-02-19"
output:
  html_document: default
---
##Project 1 for DS 6306
# We are tasked with conducting a EDA for the Beers and Breweries Data Set that contains 2410 craft beers and 558 Breweries respectivily.
# Data documents
#
#importing the datasets and analyzing the data types
```{r}
beers = read.csv("D:/Github/MSDS_6306_Doing-Data-Science/Unit 8 and 9 Case Study 1/Beers.csv")
breweries = read.csv("D:/Github/MSDS_6306_Doing-Data-Science/Unit 8 and 9 Case Study 1/Breweries.csv")
summary(beers)
summary(breweries)
unique(beers$Style)
```

##Question 1 
#how many breweries are present in each state
```{r}

library(ggplot2)
library(ggthemes)
library(tidyverse)
breweries %>% group_by(State) %>%
  summarise(count = n_distinct(Brew_ID)) %>%
  arrange(count) %>%
  ggplot(aes(x= reorder(State, +count), y = count)) +
  geom_bar(stat="identity", fill ="blue") +
  scale_x_discrete(guide = guide_axis(n.dodge=2)) + 
  geom_text(aes(label=count), vjust=-0.5, color="black",
            position = position_dodge(0.9), size=2.5) +
  ylab('# of Breweries') +
  xlab('States') +
  ggtitle('# of Breweries per State') +
  theme_pander()
  

```
#Merging the beer and brewery datasets, both share the Brew/Brewery ID column and contain no missing values so they can be inner joined. 
```{r}
library(dplyr)
#first change one of the datasets column name to match the other
breweries <- breweries %>% rename("Brewery_id"= "Brew_ID")
#Change Name of brewery to avoid duplicate column names
breweries <- breweries %>% rename("Brewery" = "Name")
data <- beers %>% inner_join(breweries, by = "Brewery_id")
#printing of the first 6 rows
head(data) #first 6 rows
tail(data) # last 6 rows
summary(data)
#checking string columns for missing cell strings
print(colSums(data == ""))

```

##Address Missing Values
#Missing data per column 
* Name : 0
* Beer_ID : 0
* ABV : 62
* Style : 5
* IBU : 1005
* Name : 0
* Ounces : 0 
* Brewery : 0
* City : 0 
* State : 0
#There was alot of data in the beer dataset, to alivate the missing data each was searched by hand using the websites untapped.com and beeradvocate.com
# if data was found it was filled in using this method here is the number of missing data after search
#Missing data per column 
* Name : 0
* Beer_ID : 0
* ABV : 18
* Style : 3
* IBU : 514
* Name : 0
* Ounces : 0 
* Brewery : 0
* City : 0 
* State : 0
```{r}

#commiting the file to csv 
write.csv(data, "D:/documents/MSDSDoingDataScience/unfilteredData.csv")
#after filling in mising data bringing back in now filtered data frame
beer_data = read.csv("D:/documents/MSDSDoingDataScience/filteredData.csv")
#summary after hand searching missing data
summary(beer_data)
print(colSums(beer_data == ""))
# checking for duplicated beers
dup = beer_data[duplicated(beer_data$Name), ]
dup
sprintf("There are a total of %d duplicated rows consisting of %d different beers", sum(duplicated(beer_data$Name)),length(unique(dup$Name))) 
     
``` 
# To remove these duplicated beers we will run the following function
```{r}
beer_data <- beer_data[!duplicated(beer_data$Name),]
beer_data
sprintf("There are now only %d beers in the dataframe", length(beer_data$Name))
```

##4.	Compute the median alcohol content and international bitterness unit for each state. Plot a bar chart to compare.
```{r}
library(ggplot2)
library(ggpubr)
library(ggthemes)
options(repr.plot.width = 10, repr.plot.height =4)

abv <- beer_data %>%
  filter(!is.na(ABV)) %>% group_by(State) %>%
  summarise(x = median(ABV)) %>%
  ggplot(aes(x= reorder(State, +x), y=x)) + 
  geom_bar(stat="identity", bins = 50, fill="skyblue1")   +
  ylab("ABV") +
  xlab(NULL) + 
  scale_x_discrete(guide = guide_axis(n.dodge=2)) +
  theme_tufte()
ibu <- beer_data %>%
  filter(!is.na(IBU)) %>% group_by(State) %>%
  summarise(x = median(IBU)) %>%
  ggplot(aes(x=reorder(State, +x), y=x)) + 
  geom_bar(stat="identity", bins = 50, fill="skyblue1")  +
  xlab("State") +
  ylab("IBU") +
  scale_x_discrete(guide = guide_axis(n.dodge=2)) +
  
  theme_tufte()
  
#combining plots together
p <- ggarrange(abv, ibu, ncol = 1, nrow = 2, align = "v") 

annotate_figure(p, top = text_grob("Median ABV and IBU per State", face ="bold", size=16, family="serif")) + theme_tufte()
```
##Which state has the maximum alcoholic (ABV) beer? Which state has the most bitter (IBU) beer?
```{r}
#options(repr.plot.width = 10, repr.plot.height =4)

beer_data %>%
  filter(!is.na(ABV)) %>% group_by(State) %>%
  summarise(x = max(ABV)) %>%
  top_n(3, x) %>%
  ggplot(aes(x= State, y=x)) + 
  geom_bar(stat="identity", fill="skyblue1")   +
  ylab("ABV") +
  xlab(NULL) +
  ggtitle('Top 3 States With Highest ABV') +
  theme_tufte()
beer_data %>%
  filter(!is.na(IBU)) %>% group_by(State) %>%
  summarise(x = max(IBU)) %>%
  top_n(3, x) %>%
  ggplot(aes(x=State, y=x)) + 
  geom_bar(stat="identity", fill="skyblue1")  +
  ylab("IBU") +
  ggtitle('Top 3 States With Highest IBU') +
  theme_tufte()
beer_data %>%
  filter(!is.na(ABV)) %>% group_by(State) %>%
  summarise(x = max(ABV)) %>% top_n(3,x)
```
##Comment on the summary statistics and distribution of the ABV variable.
```{r}
library(ggplot2)
library(dplyr)
library(hrbrthemes)
library(viridis)


beer_data %>% filter(!is.na(ABV)) %>%
  ggplot(aes(x=ABV)) + 
  geom_boxplot(color="black") +
  geom_vline(aes(xintercept= mean(ABV)), color ="red") +
  geom_text(aes(mean(ABV), 0, label=paste("Mean :",round(mean(ABV), digits= 4)), vjust = -12.7, hjust = -1.99)) +
  geom_text(aes(mean(ABV), 0, label=paste("SD :",round(sd(ABV), digits= 4)), vjust = -11.5, hjust = -2.34)) +
  geom_text(aes(mean(ABV), 0, label=paste("Min :",round(min(ABV), digits= 4)), vjust = -10.3, hjust = -2.57)) +
  geom_text(aes(mean(ABV), 0, label=paste("Max :",round(max(ABV), digits= 4)), vjust = -9.2, hjust = -2.44)) +
  geom_text(aes(mean(ABV), 0, label=paste("Med :",round(median(ABV), digits= 4)), vjust = -8.1, hjust = -2.17)) +
  ggtitle("Distribution of ABV") +
  ylab(NULL) +
  theme_stata()
  
```
##Is there an apparent relationship between the bitterness of the beer and its alcoholic content? Draw a scatter plot.  Make your best judgment of a relationship and EXPLAIN your answer.
#after conducting the visualization a non parmeteric kendall tau correlation test was used. It found that there is a moderate positive relationship between ABV vs. IBU and this value was found to be statistically significant p<0.0001.
```{r}
beer_data %>% filter(!is.na(ABV) & !is.na(IBU)) %>%
  ggplot(aes(x= ABV, y= IBU)) + 
  geom_point() +
  geom_smooth(method=lm, se=FALSE) +
  ggtitle(paste0("ABV vs. IBU relationship with coorleation cofficent of ", round(cor.test(beer_data$ABV,beer_data$IBU, method="pearson")$estimate, digits = 4), "\n and a pvalue : ", ... = cor.test(beer_data$ABV,beer_data$IBU, method="pearson")$p.value) ) +
  theme_tufte()
cor.test(beer_data$ABV, beer_data$IBU)[3] 
```
##8.	Budweiser would also like to investigate the difference with respect to IBU and ABV between IPAs (India Pale Ales) and other types of Ale (any beer with “Ale” in its name other than IPA).  You decide to use KNN classification to investigate this relationship.  Provide statistical evidence one way or the other. You can of course assume your audience is comfortable with percentages … KNN is very easy to understand conceptually. 
In addition, while you have decided to use KNN to investigate this relationship (KNN is required) you may also feel free to supplement your response to this question with any other methods or techniques you have learned.  Creativity and alternative solutions are always encouraged.  

```{r}
library(class)
library(e1071)
library(caret)
library(ggplot2)
library(dplyr)
# filtering for rows with non missing values in ABV and IBU that contain the substring 'IPA' or 'Ale', from there it is encoded to a new 
# column of type containing either the string 'IPA' or 'Ale'
beer_knn_dataset = beer_data %>%
  filter(!is.na(ABV) & !is.na(IBU) & (grepl('Ale', Style) | grepl('IPA', Style))) %>%
  mutate(Type = ifelse(!grepl('Ale', Style) ,'IPA', 'Ale'))

#creating dataframes for metrics
accs = data.frame(accuracy = numeric(25), k = numeric(25))
sens = data.frame(sensitivity = numeric(25), k = numeric(25))
spec = data.frame(specificity = numeric(25), k = numeric(25))
#Figuring out which K value to us 
for(i in 1:25)
{
  #Knn cross validation model 
  classifications = knn.cv(beer_knn_dataset[,c(3,5)],beer_knn_dataset$Type, prob = TRUE, k = i, use.all = FALSE)
  #creating a table
  table(beer_knn_dataset$Type,classifications)
  #Confusion Matrix
  CM = confusionMatrix(table(beer_knn_dataset$Type,classifications))
  #Adding the metrics to their perspective dataframes
  accs$accuracy[i] = CM$overall[1]
  sens$sensitivity[i] = CM$byClass[1]
  spec$specificity[i] = CM$byClass[2]
  #adding k value to dataframes
  accs$k[i] = i
  sens$k[i] = i
  spec$k[i] = i
}
#Plotting the metrics
ggplot() +
  geom_line(data = accs, aes(k,accuracy,  colour ="Accuracy")) +
  geom_line(data = sens ,aes(k,sensitivity,  colour ="Sensitivity")) +
  geom_line(data = spec, aes(k,specificity, colour = "Specificity")) + 
  ggtitle("ABV & IBU KNN MODEL") +
  ylab("ratio") +
  xlab("K") +
  scale_color_manual(values = c("Accuracy" = "blue", "Sensitivity" = "red", "Specificity" = "purple")) +
  labs(color = "Metric")

#KNN Cross validation model 
classifications = knn.cv(beer_knn_dataset[,c(3,5)],beer_knn_dataset$Type, prob = TRUE, k = 5, use.all=FALSE)
#Creating a Table
table(beer_knn_dataset$Type,classifications)
#Confusion Matrix 
CM = confusionMatrix(table(beer_knn_dataset$Type,classifications))
CM

```


```{r}
number_beers = beer_data %>% count(State)
number_beers %>% ggplot(aes(x=reorder(State, +n), y=n)) +
  geom_bar(stat = "identity") +
  xlab("State") +
  ylab("Count") +
  ggtitle("Number of Different Beers in Each State") +
  scale_x_discrete(guide = guide_axis(n.dodge=2)) +
  geom_text(aes(label=n), vjust=-0.5, color="black",
            position = position_dodge(0.9), size=2.5) +
  theme_tufte()
```

```{r}
top_beers =distinct(beer_data %>% group_by(State) %>% top_n(1, Style) %>% select(State, Style), State, .keep_all =TRUE)

top_beers <- top_beers %>% group_by(Style) %>% count(Style)
top_beers
top_beers$Style[1] = "Märzen / Oktoberfest"
top_beers$Style[2] = "Kölsch"
top_beers %>% ggplot(aes(y=reorder(Style, +n), x=n)) + geom_bar(stat="identity") + geom_text(aes(label=n), color="white",
          hjust=1.1, size=5) +
  ylab("Style") +
  xlab("Count") + 
  ggtitle("Count of State's Most Abundant Style of Beer") + 
  theme_tufte()
```
```{r}
# take the count of each group of ounces
temp = beer_data %>% count(Ounces)
# arrange in descending order
temp <- temp %>% arrange(desc(Ounces))
#make rows with small numbers into one group
temp$Ounces[temp$n <= 10] <- "Other"
#making ounces a character string
temp$Ounces <- as.character(temp$Ounces)
#creating a percentage column 
temp$labels <- round(temp$n / sum(temp$n) *100, digits =3 )
#combining other rows
temp = temp %>% group_by(Ounces) %>% summarize(labels = sum(labels)) %>% ungroup()
#concatentating ounces and labels column 
temp$type <- paste(temp$Ounces, " (", temp$labels, "%)" )
temp %>% ggplot(aes(x="", y=labels, fill = type)) + 
  geom_bar(stat= "identity", width=1) +
  coord_polar("y", start=0) +
  scale_fill_viridis(discrete = TRUE, name= "Ounces (Percentage)") +
  ylab(NULL) +
  xlab(NULL) +
  ggtitle("Distribution of Can Sizes") +
  theme(axis.text = element_blank(), 
        axis.ticks = element_blank(), 
        panel.grid = element_blank()) 
```
